---
title: "Product of Independent Standard Normals"
author: "Joonsuk Kang"
date: "2021-01-11"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Notation

Define $Y_k=\prod_{i=1}^k X_i$ where $X_i\sim^{i.i.d.} N(0,1)$.

Maybe we could use notation such as $Y_k\sim^{i.i.d.}N^k(0,1)$?



# Empirical distributions

```{r}
Y1 <- rnorm(n=100000, mean=0,sd=1)
Y2 <- rnorm(n=100000, mean=0,sd=1)*Y1
Y3 <- rnorm(n=100000, mean=0,sd=1)*Y2
Y4 <- rnorm(n=100000, mean=0,sd=1)*Y3
Y5 <- rnorm(n=100000, mean=0,sd=1)*Y4
```

## Empirical CDF
```{r}
plot(ecdf(Y1), main="Empirical CDF"); 
lines(ecdf(Y2), col='red')
lines(ecdf(Y3), col='blue')
lines(ecdf(Y4), col='green')
lines(ecdf(Y5), col='yellow')
```

## (Gaussian kernel) Smoothed Density

```{r}
plot(density(Y3), col='blue', main='smoothed density')
lines(density(Y2), col='red')
lines(density(Y1), col='black')
```

Notice the scale difference in y-axis

```{r}
plot(density(Y5), col='yellow', main='smoothed density')
lines(density(Y4), col='green')
lines(density(Y3), col='blue')
```


# Theoretical Results

## Moments

\[
E[Y_k^p]=E[(X_1\times X_2 \times \cdots \times X_k)^p]=(E[X_1^p])^k=((p-1)!!)^k \quad\text{if p is even;} \quad\text{ =0 if p is odd}
\]
where $(p-1)!!=(p-1)(p-3)\cdots 3\times 1$ is the double factorial.

Note that $E[Y_k]=0$ and $var(Y_k)=E[Y_k^2]=1$ for all $k$. With mean and variance fixed, the distribution in general becomes more concentrated around zero, while the tail becomes heavier as $k$ increases.



## PDF

When k=1, $Y_1$ follows the standard normal distribution:
\[
f_{Y_1}(y)=\frac{\exp(-y^2/2)}{\sqrt{2\pi}}
\]

When k=2:
\[
f_{Y_2}(y)=\frac{K_0\left( |y|\right)}{\pi}
\]
where $K_n(z)$ is a modified Bessel function of the second kind.

When k=3:
\[
f_{Y_3}(y)=\frac{1}{(\sqrt{2\pi})^3}G_{0,3}^{3,0}\left(\frac{y^2}{8}; 0,0,0 \right)
\]
where $G^{m,n}_{p,q}\left(x;^{a_1,\cdots,a_p}_{b_1,\cdots,b_q}  \right)$ is a Meijer G-function.


Reference for k=2 and k=3: https://mathworld.wolfram.com/NormalProductDistribution.html


## Characteristic Function

When k=1:
\[
\varphi_{Y_1}(t)=E[\exp(itY_1)]=\exp(-t^2/2)
\]

When k=2:
\[
\varphi_{Y_2}(t)=E[\exp(itY_2)]=\sqrt{\frac{1}{1+t^2}}
\]
which is trivial to obtain using https://en.wikipedia.org/wiki/Product_distribution#Characteristic_function_of_product_of_random_variables.


## General form of CDF

\[
P(Y_k\leq t)=1-\frac{1}{2\pi^{k/2}}G^{0,k+1}_{k+1,1}\left( \frac{2^k}{t^2}; ^{1, 1/2, \cdots, 1/2} _0 \right)
\]

From Proposition 2 in https://arxiv.org/pdf/1711.10516.pdf





## More papers

- Exact Distribution for the Product of Two Correlated
Gaussian Random Variables
(https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=7579552&casa_token=PwS246oKBVEAAAAA:GTHT402O2YFxiLJzaS_KsIfgnnzp9aWcoPf-WTlbsnoi1yTgcbS53UVp1QPD1PB4eH3ar-g7GA)




